---
title: "IRWLS algorithm for MLE in logistic regression"
subtitle: Burn injuries data
output:
  html_document:
    df_print: paged
---


### Logistic regression using our own IRWLS algorithm

An auxiliary function: 
```{r}
p.from.beta <- function(beta,x){
  lin.term <- beta[1] + x %*% beta[-1]
  e.lt <- exp(lin.term)
  p <- e.lt/(1+e.lt)
  return(list(p=p,lt=lin.term))
}
```

The main function: 
```{r}
logistic.IRWLS <- function(x,y,weights.out=1,x.new=x,
                           max.iter=10,eps.beta=1e-5,
                           plts=FALSE){
  if (plts&(dim(as.matrix(x))[2]>1)){
    plts<-FALSE
    warning("Plots are valid only when dim(x)[2]==1")
  }
  # Step 0
  stop.rule <- FALSE
  iter <- 0
  beta.0 <- coef(lm(y~x)) 
  
  while (!stop.rule){
    iter <- iter + 1 
    # step 1
    p.lt <- p.from.beta(beta.0,x)
    p <- p.lt$p
    lt <- p.lt$lt
    ps.e <- (y-p)/(p*(1-p))
    z <- lt + ps.e 
    wt <- p*(1-p) *weights.out
    
    if (plts){
      op<-par(mfrow=c(1,2))
      plot(x,y,cex=8*wt)
      lines(x,p,col=2)
      plot(x,z,cex=8*wt)
      lines(x,lt,col=2)
      par(op)
    }
    
    lm.1 <- lm(z~x,weights = wt) 
    beta.1 <- coef(lm.1)
    
    # checking stop rules
    if ((iter>=max.iter)|(sum((beta.1-beta.0)^2)<eps.beta)){
      stop.rule<-TRUE
    } else {
      beta.0 <- beta.1
    }
  }
  
  aux <- summary(lm.1)
  p.lt <- p.from.beta(beta.1,x)
  p <- p.lt$p
  lt <- p.lt$lt
  se.beta <- diag(aux$cov.unscaled)^.5
  null.devi <- sum(-2*dbinom(y,1,mean(y),log=TRUE))
  resid.devi <- sum(-2*dbinom(y,1,p,log=TRUE))
  
  return(list(coefficients=beta.1, se.coef=se.beta,
              fitted.values=p,linear.predictors=lt,
              predicted.values=p.from.beta(beta.1,x.new)$p,
              null.deviance=null.devi,
              residual.deviance=resid.devi)
         )
} 
```

### Logistic regression for Burn Injuries Data

The *Burn injuries dataset* (Fan, J. and I. Gijbels (1996). *Local polynomial modelling and its applications*) contains data from 435 adults (between ages 17 and 85) suffering from burn injuries.
It includes two variables:

* `surv`: The binary response variable is taken to be 1 for those victims who survived their burn injuries and zero otherwise.
* `lgae`: log(area of third degree burn + 1) is taken as a covariate.

The conditional expectation of `surv` given a level of `lgae` is the conditional probability of 
	survival given this particular value of `lgae`.

Apply the previous function to the estimation of the parameters of the logistic regression in the  *burn injuries data dataset* (file `burn.cont.dat`).

Reeding the data from a file:
```{r}
burns <-read.table("burn.cont.dat", head=T, sep=" ")
lgae <- burns[,2]
perm <- order(lgae)
lgae <- lgae[perm]
superv <- burns[perm,3]
```

Logistic regression using our own IRWLS algorithm:
```{r} 
x <- 
y <- superv

max.iter <- 10
eps.beta <- 1e-5 
plts <- TRUE

our.glm.burn <- logistic.IRWLS(x=matrix(lgae,ncol=1), y=superv, 
               weights.out=1, max.iter=10,
               eps.beta=1e-5, plts=TRUE)

# names(our.glm.burn)
# [1] "coefficients"      "se.coef"           "fitted.values"     "linear.predictors"
# [5] "predicted.values"  "null.deviance"     "residual.deviance"
``` 
Coefficient | Estimate | Std. Error
----------- | -------- | -----------
`r names(our.glm.burn$coefficients)[1]`|`r our.glm.burn$coefficients[1]`|`r our.glm.burn$se.coef[1]`
`r names(our.glm.burn$coefficients)[2]`|`r our.glm.burn$coefficients[2]`|`r our.glm.burn$se.coef[2]`

Null Deviance:      `r our.glm.burn$null.deviance`

Residual Deviance:  `r our.glm.burn$residual.deviance`

### Comparing with `glm` of R

Compare your result with that obtained with the function `glm` of R.
How much of the information you get when doing the `summary` of the object produced by `glm` could you obtain from your algorithm?

Fitting the logistic regresion using glm
```{r}
glm.burn <- glm(superv~lgae,family=binomial())
summary(glm.burn)
```

Comparing glm results and our own results:
```{r}
plot(lgae,superv,main="Burn injuries data. Logistic regression")
lines(lgae,glm.burn$fitted.values,col=4)
lines(lgae,our.glm.burn$fitted.values,col=2, lty=2)
legend("bottomleft",c("glm","our glm"),col=c(4,2),lty=c(1,2))
```
